The Influence of Radial Drift and Gas Accretion on the C/O Ratio in Protoplanetary Disks

### 2016-04-15 Dissertate-Harvard-LaTeX/chapters/chapter3.tex
upd

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{figures/drift_timescales_betaS1_gas_acc_new2.pdf}
%\vspace{-0.5in}
\caption{Relevant timescales for dynamical effects in the desorption process: $t_{\rm drift}$ (solid lines), $t_{\rm des}$ (dashed lines) and $t_{\rm gas, acc}$ (dotted lines). The timescales are calculated at three representative locations, i.e. the H$_2$O, CO$_2$ and CO snowlines in the static disk. For our choice of parameters, the snowlines are located at $\sim$0.7 AU (blue lines), $\sim$8.6 AU (green lines) and $\sim$59 AU (red lines), respectively. The horizontal dot-dashed line represents a typical disk lifetime of 3 Myr. The particle size ordering at the minimum $t_{\rm drift}$ is not monotonic in snowline distance due to different drag regimes for those particle sizes at the snowline locations (Epstein drag at the H$_2$O and CO$_2$ snowlines, and Stokes drag at the CO snowline). Similarly, the ordering of $t_{\rm des}$ is not monotonic in snowline distance due to the non-monotony in mean molecular weight between H$_2$O, CO$_2$ and CO (18 $m_{\rm p}$, 44 $m_{\rm p}$ and 28 $m_{\rm p}$, respectively). Radial drift and gas accretion affect desorption in the regions where their respective timescales, i.e. $t_{\rm drift}$ and $t_{\rm gas, acc}$, are comparable to the desorption timescale $t_{\rm des}$.} 
\label{fig:timescales}
\end{figure}

@@

\begin{figure*}[tb]
\centering
\includegraphics[width=0.7\textwidth]{figures/desorption_distance_passive_active_colorbar_test2.pdf}
%\vspace{-0.5in}
\caption{Desorption distance (if a grain fully desorbs; horizontal lines) or final distance (if a grain does not fully desorb; diagonal lines for particles that do not drift and non-horizontal, non-diagonal lines for particles that drift), as a function of a particle's initial location in the disk, for a range of particle sizes, and for both an irradiated disk (left panels) and an evolving disk (right panels). The desorption distance is calculated for particles composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels). The desorption distance for a static disk is shown for comparison (dashed vertical and horizontal lines). The particle size increases from $10^{-3}$ cm to $10^8$ cm as indicated by the color bar. For a particle of a given initial size that entirely desorbs during $t_{\rm d}=3$ Myr, the desorption distance is the same regardless of the particle's initial location.} 
\label{fig:snowlines}
\end{figure*}

\begin{figure*}[t!]
\centering
\includegraphics[width=0.7\textwidth]{figures/desorption_distance_actual_vs_estimated_passive_active_new.pdf}
%\vspace{-0.5in}
\caption{Desorption distance estimated from analytic calculations (see text) as a function of the desorption distance calculated numerically, for the range of particle sizes that desorb at a fixed distance regardless of their initial location (see Figure \ref{fig:snowlines} and text). The estimate is performed for an irradiated disk (left panels) and an evolving disk (right panels).  The particles are composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels). The analytic approximation is in good agreement with the numerical result for most cases, with the exception of larger particles, $s \gtrsim 10$ cm (see text).}
\label{fig:an_vs_actual}
\end{figure*}

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=0.8\textwidth]{figures/s_t_a.pdf}
%\vspace{-0.5in}
\caption{Left panels: size of desorbing H$_2$O particles as a function of time, for different initial particle sizes and for three initial locations in an irradiated disk: 1 AU (top left), 10 AU (middle left) and 100 AU (bottom left). Particles desorb almost instantaneously. Right panel: fractional mass of the desorbing particles as a function of the particle's location as it drifts, for different initial particle sizes, and at the same initial locations presented in the left panel. Particles lose most of their mass very close to the distance at which they fully desorb. The static H$_2$O snowline is shown for reference (dashed vertical lines).}
\label{fig:s_t_a}
\end{figure*}

@@

@@

Given $r_{\rm des}$, we need to only calculate the distance over which particles desorb to determine the location of a snowline. Figure \ref{fig:s_t_a}, left panels, shows the size evolution with time for H$_2$O particles of various initial sizes, starting at three different initial locations in an irradiated disk. Once solid H$_2$O particles begin to evaporate, they do so almost instantly for all explored particle sizes and initial locations. %, although the time $t_{\rm des}$ at which a particle of a given initial size desorbs depends on its initial distance. 
%A particle located at the initial time $t_0$ at a distance such that $t_{\rm des}>t_{\rm d}$ will therefore not desorb during the disk lifetime, but our model assumes that particles drift continuously at any location in the disk. [not sure you need to explain this]
The right panels of Figure \ref{fig:s_t_a} show that the drifting grains lose most of their mass in a very narrow distance range; moreover, this distance is the same for a given initial particle size, no matter where the particle started drifting at the time $t_0$ when the simulation is started. Figure \ref{fig:s_t_a} thus demonstrates that solid particles that drift and fully desorb during the lifetime of the protoplanetary disk do so (1) instantaneously, and (2) at a fixed stellocentric distance, regardless of their initial location in the disk. 
It follows that the H$_2$O, CO$_2$ and CO snowlines are fixed for a given initial particle size and disk model. Both of these conclusions remain valid for an evolving and a viscous disk, as well as for particles composed of CO$_2$ or CO, but the snowline locations will vary between the three disks for a given initial particle size (see Section \ref{sec:COratio}). If we do not take into account the time dependence of the mass accretion rate and stellar luminosity (see Section \ref{sec:disk}), the C/O ratio will then only depend on disk properties, grain size, and the abundance of H$_2$O, CO$_2$ and CO relative to the H$_2$ abundance in the disk midplane, and {\it not} directly on the disk age when only considering drift, accretion and desorption.  %\emgr{I've yet to figure out why there is a discrepancy for the small grains. I'll keep thinking about it, but for now I'll leave it as it is since I don't want to give a false explanation just to have one.}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=0.55\textwidth]{figures/C_O_ratio_passive_active_disk_many_colorbar_complete_new2.pdf}
%\vspace{-0.5in}
\caption{Estimated C/O ratio in gas (solid lines) and in dust (dashed lines) for an irradiated disk (top panel), an evolving disk (middle panel) and a viscous disk (bottom panel).
%, for the range of particle sizes that desorb at a fixed distance regardless of their initial location in the disk. 
The particle size increases from $\sim$0.05 cm to $\sim$700 cm as indicated by the color bar. The horizontal dotted line represents the stellar value of 0.54. The black lines represent the C/O ratio in gas (solid black line) and dust (dashed black line) for a static disk, with the temperature profile given by Equation (\ref{eq:diskT}) for the top two panels and by Equation (\ref{eq:activeT}) for the bottom panel. 
%The snowline location moves inward as the particle size increases. 
%The curves exterior to the static snowline represent desorption distances in the absence of readsorption.  %Taking  readsorption into account, the snowlines for these particle sizes are in fact coincident with the static snowline. 
For both the evolving and the viscous disk, the movement of desorbed CO$_2$ gas inside the CO$_2$ snowline, and of desorbed CO$_2$ and H$_2$O gas inside the H$_2$O snowline due to gas accretion will increase the amount of oxygen gas inside the respective snowlines and thus reduce the gas C/O ratio, as shown by the arrows.}
%, with the decrease more significant interior to the H$_2$O snowline.}
\label{fig:CO_ratio}
\end{figure}

Given our results in Section \ref{sec:snowlines}, a disk's C/O ratio in mainly affected by the snowline location for the particle size housing the most mass in ice.
%We now use our model and the results of Section \ref{sec:snowlines} to determine the H$_2$O, CO$_2$ and CO snowline locations in disks with static chemistry that experience radial drift of solids and gas accretion onto the central star. Using this and assuming a fixed abundance relative to hydrogen for each of these volatiles, we can then estimate the C/O ratio throughout the disk as a function of the particle size. 
Realistic grain size distributions in disks are dominated by large grains (e.g., \citealt{dalessio01}, \citealt{birnstiel12}). In Figure \ref{fig:CO_ratio}, we display the H$_2$O, CO$_2$, and CO snowline locations as a function of particle size for disks with static chemistry that experience radial drift of solids and gas accretion onto the central star.  The minimum snowline distance for a disk is given by the curve corresponding to the maximum particle size it hosts.  For grains that have grown to radii larger than $\sim$7 m and are able to drift and desorb, the $\sim$7 m snowline applies (see Section 3).  %Thus for a given grain size distribution with a maximum particle size, we can pick out the appropriate minimum snowline locations from Figures \ref{fig:snowlines} and \ref{fig:CO_ratio}, as we show later in this section.

@@

Figure \ref{fig:CO_ratio} shows the estimated C/O ratio in gas and dust as a function of semimajor axis for an irradiated disk, an evolving disk, and a viscous disk, under the simplifying assumption that the abundance relative to hydrogen for each volatile is fixed, so that drift and accretion affect only the locations of the snowlines. We use the relative number densities of C and O in their different molecular forms (H$_2$O, CO$_2$ and CO) from Table 1 of \citet{oberg11}.  Snowline locations correspond to $r_{\rm des}$ in Figure \ref{fig:snowlines}, representing the location at which particles desorb in the absence of readsorption.  The C/O ratio for a static disk, where desorption and readsorption balance \citep{hollenbach09}, is shown as a guideline. We note that the true snowline for particles with $r_{\rm des}$ outside the static snowline is the static snowline itself---thus only particles with initial sizes larger than $\sim$0.05 cm are plotted in the three panels, as particles that form snowlines at larger distances (cf. Figure \ref{fig:snowlines}) are not true snowlines.

Before discussing the quantitative aspects of this plot, it is essential to acknowledge that our estimates for the C/O ratios in the evolving and viscous disks ignore the movement of the desorbed ices with the accreting gas---the relative fluxes of the volatiles in gaseous and solid form will affect the relative abundance of C and O in gas and dust throughout the disk. As demonstrated in Figure \ref{fig:s_t_a}, this will not affect the snowline locations for particles of a given size, but will change the shape of the C/O curves in between the various snowlines. For example, for the disk parameters and particle sizes displayed in Figure \ref{fig:CO_ratio}, water molecules in solid particles drift up to $\sim$1000 times faster across the H$_2$O snowline than do molecules of CO and CO$_2$ vapor that are entrained in the accreting gas.  This differential inward motion will result in an increased oxygen gas abundance inside the H$_2$O snowline, and thus a (in some cases much) lower gaseous C/O ratio in this region. Conversely, oxygen gas inside the water snowline will be depleted compared to the static disk if H$_2$O particles grow to planetesimal size and stall their migration between the H$_2$O and CO$_2$ snowlines, leaving only gaseous CO and CO$_2$ to accrete inward.  Growth of large planetesimals can therefore increase the C/O ratio in the inner disk.

@@

Figure \ref{fig:CO_ratio} plots snowline curves for particle sizes $\sim0.5$ cm $\lesssim $s$ \lesssim$ 7m.  In the outermost disk, H$_2$O, CO$_2$, and CO all solidify.  Hence, relative drift across the CO snowline can alter only the abundances of volatiles between the CO$_2$ and CO snowlines, but not the C/O ratio in this region.  Interior to the CO$_2$ snowline, however, relative drift is important. 
We have found that the largest drifting particles in our model ($\sim$7 m) drift faster than the gas at both the H$_2$O and CO$_2$ snowlines. We thus conclude that the C/O ratio interior to the H$_2$O and CO$_2$ snowlines in our evolving and viscous disks will be lower than in the static disk, due to the additional oxygen added to the gas by desorbing H$_2$O and CO$_2$.  For these particle sizes, our calculated C/O ratio is an upper limit,  as indicated by the arrows in Figure \ref{fig:CO_ratio}.

@@

As noted earlier, we assume that the total (ice and gas) abundance of each volatile is the same at every radius after ices have migrated. This is a good approximation for the irradiated disk, given that this model, by definition, presents a constant influx of particles at any given radius while the gas is static, and thus the ice and gas surface density remain constant. For the evolving disk, the gas-phase C/O ratio may decrease everywhere interior to the H$_2$O and CO$_2$ snowlines due to the decline in the surface density of solids with time at any given radius. For the viscous disk, the solid abundances at a fixed radius are constant, given that this model is not time-dependent, but the gas-to-solid ratio is not constant, which can result in a substantially lower C/O ratio interior to the H$_2$O and CO$_2$ snowlines compared to the static case (as indicated by the arrows in Figure \ref{fig:CO_ratio}). The main goal of Figure \ref{fig:CO_ratio}, however, is to show the different snowline radii in static and non-static disks, and therefore the locations in the disk where the gas-phase C/O ratio is reduced or increased, rather than provide a quantitative estimate of the magnitude of the C/O increase or decrease.

@@

Perhaps the most interesting feature is the fact that the snowlines are pushed inwards as the grain size increases. While the plot only shows the snowlines and C/O ratio for particle sizes up to $\sim$7 m, we have found that %almost kilometer-sized boulders are able to drift and desorb for both the passive and the active disk. B
bodies larger than $\sim$7 m evaporate at the same location as the meter-sized planetesimals (see Section \ref{sec:snowlines}). However, the contribution of kilometer-sized bodies to the snowline location is modest, since they only drift if they are located very close to the snowline.
%[new sentence explaining that their contribution may not be very big due to their modest drift.] though 
%only the kilometer-sized boulders that are very close to the snownlines will contribute to determining the snowline location. 
Thus the innermost snowlines (depicted in red in Figure \ref{fig:CO_ratio}) set the limit on how close in the H$_2$O, CO$_2$ and CO snowlines can be pushed due to radial drift and gas accretion on to the host star. For a grain size distribution with a maximum particle size different than our model, one can pick out the appropriate minimum snowline locations from this plot.%Realistic grain size distributions in disks are dominated by large grains (e.g., \citealt{dalessio01}, \citealt{birnstiel12}).  Thus for a given grain size distribution with a maximum particle size, we can pick out the appropriate minimum snowline locations from this plot. %Our model thus sets an inner limit to the snowline locations. %Therefore, the snowlines produced by the largest drifting solids in our model set the inner limit on the snowline locations and are insensitive to a particular grain size distribution --- for a given grain size .

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=\textwidth]{figures/time_plots.pdf}
%\vspace{-0.5in}
\caption{Desorption or final distance as a function of initial position in the disk for particles of initial size $s_0=1$ m (left panels) and $s_0=10$ cm (right panels), for grains composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels). The evolution is shown at four representative timescales: $10^4$ yr (cyan curve), $10^5$ yr (purple curve), 1 Myr (blue curve), and 3 Myr, the disk lifetime (orange curve). For a given particle size, the desorption distance, and hence the H$_2$O, CO$_2$ and CO snowlines, have the same location regardless of the time at which the simulation is stopped.}
\label{fig:timeplots}
\end{figure*}

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{figures/desorption_distance_varying_Md.pdf}
%\vspace{-0.5in}
\caption{Desorption or final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for total disk masses $M=0.1 M_{\odot}$ (solid lines) and $M=0.01 M_{\odot}$ (dashed lines). The timescales at which we stop the simulations are $10^4$ yr (cyan curve), $10^5$ yr (red curve), 1 Myr (green curve) and 3 Myr (blue curve). A lower disk mass does not change the snowline location.}
\label{fig:varMd}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{figures/desorption_distance_transition_disk_1000.pdf}
%\vspace{-0.5in}
\caption{Desorption or final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for our fiducial disk (solid lines) and for a transition disk with an inner cavity at $r_0=4$ AU (dashed lines). The timescales of the simulations and their color code are the same as in Figure \ref{fig:varMd}. Particles that start inside the cavity drift towards the original snowline, while particles that start outside the gap stop shortly after crossing the gap edge, due to being trapped in a pressure maximum.}
\label{fig:cavity}
\end{figure}

@@

@@

@@

@@

@@

\begin{table}[t!]
\caption{The effects of dynamical and chemical processes on snowline shapes and locations}
\begin{center}
\begin{tabular}{|l|l|}\hline
\textbf{Process} & \textbf{Effect} \\\hline
Radial drift & $\leftarrow$ \footnote{The arrows signify how a process affects the snowline: $\leftarrow$ means that the snowline is pushed closer to the host star compared to the static snowline, $\rightarrow$ means that the snowline is pushed further from the host star compared to the static snowline. The presence of both arrows means that the process may have both effects on the snowline location.} \\\hline
Gas accretion & $\leftarrow$ \footnote{Gas accretion pushes the snowlines inwards compared to the snowline locations in a static disk. However, accretional heating may push the snowline outwards compared to an evolving disk without accretion heating.} \\\hline
Particle growth & $\leftarrow$ \footnote{As stated in the main text, if particles grow to km-sizes and above, the snowline remains the same as that of a static disk.} \\\hline
Turbulent diffusion & $\rightarrow$ $\leftarrow$ \\\hline
Particle fragmentation & $\rightarrow$ $\leftarrow$ \\\hline
Grain morphology & $\rightarrow$ \\\hline
Particle composition & $\rightarrow$ $\leftarrow$ \\\hline
Disk gaps and holes & $\rightarrow$ \\\hline
Accretion rate evolution & $\rightarrow$ $\leftarrow$ \\\hline
Stellar luminosity evolution & $\leftarrow$ \\\hline
Non-static chemistry & $\rightarrow$ $\leftarrow$ \\\hline
\end{tabular}

\end{center}
\end{table}
%causing it to cross the snowline and refreeze, thus moving the snowline location outwards altogether.

>>>>

\begin{figure}[H]
\centering
\includegraphics[width=0.7\textwidth]{figures/drift_timescales_betaS1_gas_acc_new2.pdf}
%\vspace{-0.5in}
\caption{Relevant timescales for dynamical effects in the desorption process: $t_{\rm drift}$ (solid lines), $t_{\rm des}$ (dashed lines) and $t_{\rm gas, acc}$ (dotted lines). The timescales are calculated at three representative locations, i.e. the H$_2$O, CO$_2$ and CO snowlines in the static disk. For our choice of parameters, the snowlines are located at $\sim$0.7 AU (blue lines), $\sim$8.6 AU (green lines) and $\sim$59 AU (red lines), respectively. The horizontal dot-dashed line represents a typical disk lifetime of 3 Myr. The particle size ordering at the minimum $t_{\rm drift}$ is not monotonic in snowline distance due to different drag regimes for those particle sizes at the snowline locations (Epstein drag at the H$_2$O and CO$_2$ snowlines, and Stokes drag at the CO snowline). Similarly, the ordering of $t_{\rm des}$ is not monotonic in snowline distance due to the non-monotony in mean molecular weight between H$_2$O, CO$_2$ and CO (18 $m_{\rm p}$, 44 $m_{\rm p}$ and 28 $m_{\rm p}$, respectively). Radial drift and gas accretion affect desorption in the regions where their respective timescales, i.e. $t_{\rm drift}$ and $t_{\rm gas, acc}$, are comparable to the desorption timescale $t_{\rm des}$.} 
\label{fig:timescales}
\end{figure}

@@

@@

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.8\textwidth]{figures/desorption_distance_passive_active_colorbar_test2.pdf}
%\vspace{-0.5in}
\caption{Desorption distance (if a grain fully desorbs; horizontal lines) or final distance (if a grain does not fully desorb; diagonal lines for particles that do not drift and non-horizontal, non-diagonal lines for particles that drift), as a function of a particle's initial location in the disk, for a range of particle sizes, and for both an irradiated disk (left panels) and an evolving disk (right panels). The desorption distance is calculated for particles composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels). The desorption distance for a static disk is shown for comparison (dashed vertical and horizontal lines). The particle size increases from $10^{-3}$ cm to $10^8$ cm as indicated by the color bar. For a particle of a given initial size that entirely desorbs during $t_{\rm d}=3$ Myr, the desorption distance is the same regardless of the particle's initial location.} 
\label{fig:snowlines}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/desorption_distance_actual_vs_estimated_passive_active_new.pdf}
%\vspace{-0.5in}
\caption{Desorption distance estimated from analytic calculations (see text) as a function of the desorption distance calculated numerically, for the range of particle sizes that desorb at a fixed distance regardless of their initial location (see Figure \ref{fig:snowlines} and text). The estimate is performed for an irradiated disk (left panels) and an evolving disk (right panels).  The particles are composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels). The analytic approximation is in good agreement with the numerical result for most cases, with the exception of larger particles, $s \gtrsim 10$ cm (see text).}
\label{fig:an_vs_actual}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.8\textwidth]{figures/s_t_a.pdf}
%\vspace{-0.5in}
\caption{Left panels: size of desorbing H$_2$O particles as a function of time, for different initial particle sizes and for three initial locations in an irradiated disk: 1 AU (top left), 10 AU (middle left) and 100 AU (bottom left). Particles desorb almost instantaneously. Right panel: fractional mass of the desorbing particles as a function of the particle's location as it drifts, for different initial particle sizes, and at the same initial locations presented in the left panel. Particles lose most of their mass very close to the distance at which they fully desorb. The static H$_2$O snowline is shown for reference (dashed vertical lines).}
\label{fig:s_t_a}
\end{figure}

Given $r_{\rm des}$, we need to only calculate the distance over which particles desorb to determine the location of a snowline. Figure \ref{fig:s_t_a}, left panels, shows the size evolution with time for H$_2$O particles of various initial sizes, starting at three different initial locations in an irradiated disk. Once solid H$_2$O particles begin to evaporate, they do so almost instantly for all explored particle sizes and initial locations.

The right panels of Figure \ref{fig:s_t_a} show that the drifting grains lose most of their mass in a very narrow distance range; moreover, this distance is the same for a given initial particle size, no matter where the particle started drifting at the time $t_0$ when the simulation is started. Figure \ref{fig:s_t_a} thus demonstrates that solid particles that drift and fully desorb during the lifetime of the protoplanetary disk do so (1) instantaneously, and (2) at a fixed stellocentric distance, regardless of their initial location in the disk. 
It follows that the H$_2$O, CO$_2$ and CO snowlines are fixed for a given initial particle size and disk model. Both of these conclusions remain valid for an evolving and a viscous disk, as well as for particles composed of CO$_2$ or CO, but the snowline locations will vary between the three disks for a given initial particle size (see Section \ref{sec:COratio}). If we do not take into account the time dependence of the mass accretion rate and stellar luminosity (see Section \ref{sec:disk}), the C/O ratio will then only depend on disk properties, grain size, and the abundance of H$_2$O, CO$_2$ and CO relative to the H$_2$ abundance in the disk midplane, and {\it not} directly on the disk age when only considering drift, accretion and desorption.  %\emgr{I've yet to figure out why there is a discrepancy for the small grains. I'll keep thinking about it, but for now I'll leave it as it is since I don't want to give a false explanation just to have one.}

@@

Given our results in Section \ref{sec:snowlines}, a disk's C/O ratio in mainly affected by the snowline location for the particle size housing the most mass in ice.
%We now use our model and the results of Section \ref{sec:snowlines} to determine the H$_2$O, CO$_2$ and CO snowline locations in disks with static chemistry that experience radial drift of solids and gas accretion onto the central star. Using this and assuming a fixed abundance relative to hydrogen for each of these volatiles, we can then estimate the C/O ratio throughout the disk as a function of the particle size. 
Realistic grain size distributions in disks are dominated by large grains (e.g., \citealt{dalessio01}, \citealt{birnstiel12}). In Figure \ref{1fig:CO_ratio}, we display the H$_2$O, CO$_2$, and CO snowline locations as a function of particle size for disks with static chemistry that experience radial drift of solids and gas accretion onto the central star.  The minimum snowline distance for a disk is given by the curve corresponding to the maximum particle size it hosts.  For grains that have grown to radii larger than $\sim$7 m and are able to drift and desorb, the $\sim$7 m snowline applies (see Section 3).  %Thus for a given grain size distribution with a maximum particle size, we can pick out the appropriate minimum snowline locations from Figures \ref{fig:snowlines} and \ref{fig:CO_ratio}, as we show later in this section.

\begin{figure}[H]
\centering
\includegraphics[width=0.7\textwidth]{figures/C_O_ratio_passive_active_disk_many_colorbar_complete_new2.pdf}
%\vspace{-0.5in}
\caption{Estimated C/O ratio in gas (solid lines) and in dust (dashed lines) for an irradiated disk (top panel), an evolving disk (middle panel) and a viscous disk (bottom panel).
%, for the range of particle sizes that desorb at a fixed distance regardless of their initial location in the disk. 
The particle size increases from $\sim$0.05 cm to $\sim$700 cm as indicated by the color bar. The horizontal dotted line represents the stellar value of 0.54. The black lines represent the C/O ratio in gas (solid black line) and dust (dashed black line) for a static disk, with the temperature profile given by Equation (\ref{eq:diskT}) for the top two panels and by Equation (\ref{eq:activeT}) for the bottom panel. 
%The snowline location moves inward as the particle size increases. 
%The curves exterior to the static snowline represent desorption distances in the absence of readsorption.  %Taking  readsorption into account, the snowlines for these particle sizes are in fact coincident with the static snowline. 
For both the evolving and the viscous disk, the movement of desorbed CO$_2$ gas inside the CO$_2$ snowline, and of desorbed CO$_2$ and H$_2$O gas inside the H$_2$O snowline due to gas accretion will increase the amount of oxygen gas inside the respective snowlines and thus reduce the gas C/O ratio, as shown by the arrows.}
%, with the decrease more significant interior to the H$_2$O snowline.}
\label{1fig:CO_ratio}
\end{figure}

@@

Figure \ref{1fig:CO_ratio} shows the estimated C/O ratio in gas and dust as a function of semimajor axis for an irradiated disk, an evolving disk, and a viscous disk, under the simplifying assumption that the abundance relative to hydrogen for each volatile is fixed, so that drift and accretion affect only the locations of the snowlines. We use the relative number densities of C and O in their different molecular forms (H$_2$O, CO$_2$ and CO) from Table 1 of \citet{oberg11}.  Snowline locations correspond to $r_{\rm des}$ in Figure \ref{fig:snowlines}, representing the location at which particles desorb in the absence of readsorption.  The C/O ratio for a static disk, where desorption and readsorption balance \citep{hollenbach09}, is shown as a guideline. We note that the true snowline for particles with $r_{\rm des}$ outside the static snowline is the static snowline itself---thus only particles with initial sizes larger than $\sim$0.05 cm are plotted in the three panels, as particles that form snowlines at larger distances (cf. Figure \ref{fig:snowlines}) are not true snowlines.

Before discussing the quantitative aspects of this plot, it is essential to acknowledge that our estimates for the C/O ratios in the evolving and viscous disks ignore the movement of the desorbed ices with the accreting gas---the relative fluxes of the volatiles in gaseous and solid form will affect the relative abundance of C and O in gas and dust throughout the disk. As demonstrated in Figure \ref{fig:s_t_a}, this will not affect the snowline locations for particles of a given size, but will change the shape of the C/O curves in between the various snowlines. For example, for the disk parameters and particle sizes displayed in Figure \ref{1fig:CO_ratio}, water molecules in solid particles drift up to $\sim$1000 times faster across the H$_2$O snowline than do molecules of CO and CO$_2$ vapor that are entrained in the accreting gas.  This differential inward motion will result in an increased oxygen gas abundance inside the H$_2$O snowline, and thus a (in some cases much) lower gaseous C/O ratio in this region. Conversely, oxygen gas inside the water snowline will be depleted compared to the static disk if H$_2$O particles grow to planetesimal size and stall their migration between the H$_2$O and CO$_2$ snowlines, leaving only gaseous CO and CO$_2$ to accrete inward.  Growth of large planetesimals can therefore increase the C/O ratio in the inner disk.

@@

Figure \ref{1fig:CO_ratio} plots snowline curves for particle sizes $\sim0.5$ cm $\lesssim $s$ \lesssim$ 7m.  In the outermost disk, H$_2$O, CO$_2$, and CO all solidify.  Hence, relative drift across the CO snowline can alter only the abundances of volatiles between the CO$_2$ and CO snowlines, but not the C/O ratio in this region.  Interior to the CO$_2$ snowline, however, relative drift is important. 
We have found that the largest drifting particles in our model ($\sim$7 m) drift faster than the gas at both the H$_2$O and CO$_2$ snowlines. We thus conclude that the C/O ratio interior to the H$_2$O and CO$_2$ snowlines in our evolving and viscous disks will be lower than in the static disk, due to the additional oxygen added to the gas by desorbing H$_2$O and CO$_2$.  For these particle sizes, our calculated C/O ratio is an upper limit,  as indicated by the arrows in Figure \ref{1fig:CO_ratio}.

@@

As noted earlier, we assume that the total (ice and gas) abundance of each volatile is the same at every radius after ices have migrated. This is a good approximation for the irradiated disk, given that this model, by definition, presents a constant influx of particles at any given radius while the gas is static, and thus the ice and gas surface density remain constant. For the evolving disk, the gas-phase C/O ratio may decrease everywhere interior to the H$_2$O and CO$_2$ snowlines due to the decline in the surface density of solids with time at any given radius. For the viscous disk, the solid abundances at a fixed radius are constant, given that this model is not time-dependent, but the gas-to-solid ratio is not constant, which can result in a substantially lower C/O ratio interior to the H$_2$O and CO$_2$ snowlines compared to the static case (as indicated by the arrows in Figure \ref{1fig:CO_ratio}). The main goal of Figure \ref{1fig:CO_ratio}, however, is to show the different snowline radii in static and non-static disks, and therefore the locations in the disk where the gas-phase C/O ratio is reduced or increased, rather than provide a quantitative estimate of the magnitude of the C/O increase or decrease.

@@

Perhaps the most interesting feature is the fact that the snowlines are pushed inwards as the grain size increases. While the plot only shows the snowlines and C/O ratio for particle sizes up to $\sim$7 m, we have found that %almost kilometer-sized boulders are able to drift and desorb for both the passive and the active disk. B
bodies larger than $\sim$7 m evaporate at the same location as the meter-sized planetesimals (see Section \ref{sec:snowlines}). However, the contribution of kilometer-sized bodies to the snowline location is modest, since they only drift if they are located very close to the snowline.
%[new sentence explaining that their contribution may not be very big due to their modest drift.] though 
%only the kilometer-sized boulders that are very close to the snownlines will contribute to determining the snowline location. 
Thus the innermost snowlines (depicted in red in Figure \ref{1fig:CO_ratio}) set the limit on how close in the H$_2$O, CO$_2$ and CO snowlines can be pushed due to radial drift and gas accretion on to the host star. For a grain size distribution with a maximum particle size different than our model, one can pick out the appropriate minimum snowline locations from this plot.%Realistic grain size distributions in disks are dominated by large grains (e.g., \citealt{dalessio01}, \citealt{birnstiel12}).  Thus for a given grain size distribution with a maximum particle size, we can pick out the appropriate minimum snowline locations from this plot. %Our model thus sets an inner limit to the snowline locations. %Therefore, the snowlines produced by the largest drifting solids in our model set the inner limit on the snowline locations and are insensitive to a particular grain size distribution --- for a given grain size .

@@

@@

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/time_plots.pdf}
%\vspace{-0.5in}
\caption{Desorption or final distance as a function of initial position in the disk for particles of initial size $s_0=1$ m (left panels) and $s_0=10$ cm (right panels), for grains composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels). The evolution is shown at four representative timescales: $10^4$ yr (cyan curve), $10^5$ yr (purple curve), 1 Myr (blue curve), and 3 Myr, the disk lifetime (orange curve). For a given particle size, the desorption distance, and hence the H$_2$O, CO$_2$ and CO snowlines, have the same location regardless of the time at which the simulation is stopped.}
\label{fig:timeplots}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.7\textwidth]{figures/desorption_distance_varying_Md.pdf}
%\vspace{-0.5in}
\caption{Desorption or final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for total disk masses $M=0.1 M_{\odot}$ (solid lines) and $M=0.01 M_{\odot}$ (dashed lines). The timescales at which we stop the simulations are $10^4$ yr (cyan curve), $10^5$ yr (red curve), 1 Myr (green curve) and 3 Myr (blue curve). A lower disk mass does not change the snowline location.}
\label{fig:varMd}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.7\textwidth]{figures/desorption_distance_transition_disk_1000.pdf}
%\vspace{-0.5in}
\caption{Desorption or final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for our fiducial disk (solid lines) and for a transition disk with an inner cavity at $r_0=4$ AU (dashed lines). The timescales of the simulations and their color code are the same as in Figure \ref{fig:varMd}. Particles that start inside the cavity drift towards the original snowline, while particles that start outside the gap stop shortly after crossing the gap edge, due to being trapped in a pressure maximum.}
\label{fig:cavity}
\end{figure}

@@

\begin{table}[H]
\caption{The effects of dynamical and chemical processes on snowline shapes and locations}
\begin{center}
\begin{tabular}{|l|l|}\hline
\textbf{Process} & \textbf{Effect} \\\hline
Radial drift & $\leftarrow$ \footnote{The arrows signify how a process affects the snowline: $\leftarrow$ means that the snowline is pushed closer to the host star compared to the static snowline, $\rightarrow$ means that the snowline is pushed further from the host star compared to the static snowline. The presence of both arrows means that the process may have both effects on the snowline location.} \\\hline
Gas accretion & $\leftarrow$ \footnote{Gas accretion pushes the snowlines inwards compared to the snowline locations in a static disk. However, accretional heating may push the snowline outwards compared to an evolving disk without accretion heating.} \\\hline
Particle growth & $\leftarrow$ \footnote{As stated in the main text, if particles grow to km-sizes and above, the snowline remains the same as that of a static disk.} \\\hline
Turbulent diffusion & $\rightarrow$ $\leftarrow$ \\\hline
Particle fragmentation & $\rightarrow$ $\leftarrow$ \\\hline
Grain morphology & $\rightarrow$ \\\hline
Particle composition & $\rightarrow$ $\leftarrow$ \\\hline
Disk gaps and holes & $\rightarrow$ \\\hline
Accretion rate evolution & $\rightarrow$ $\leftarrow$ \\\hline
Stellar luminosity evolution & $\leftarrow$ \\\hline
Non-static chemistry & $\rightarrow$ $\leftarrow$ \\\hline
\end{tabular}

\end{center}
\end{table}

@@

%causing it to cross the snowline and refreeze, thus moving the snowline location outwards altogether.

\item \textbf{Particle fragmentation.} Frequent particle collisions in disks cause them to fragment (e.g., \citealt{birnstiel12}). The fragmentation of meter- to km-sized particles will move the snowlines outwards, as smaller particles desorb faster and further out from the host star (cf. Figures \ref{fig:snowlines} and \ref{1fig:CO_ratio}). Large boulders, which neither drift nor desorb, may become e.g. meter-sized due to collisions and subsequent fragmentation, which will cause them to drift significantly before desorbing, pushing the snowlines inwards. Thus fragmentation can move the snowline locations in either radial direction --- specifically, fragmentation leads to a certain grain size distribution, and the largest particles in this size distribution are the ones that determine the position of the snowline.
